---
title: "Midterm Test"
author: "Steven Todd McIntyre II"
cache: true
format: 
  html:
    code-fold: true
    embed-resources: true
---

## Install Dependencies

```{r}
library(tidyverse)
library(glue)
library(here)
library(rio)
library(janitor)
```
## Part A: Data

```{r}
# create function
download_file <- function(year) {
  link <- glue::glue("https://www.oregon.gov/ode/educator-resources/assessment/TestResults20{year}/pagr_schools_ela_raceethnicity_{year-1}{year}.xlsx")
  rio::import(link, setclass = "tibble", na = c("-", "--", "*"))
}
```

### 1. (10 points)

Use the function above to download all of the data from the 2014-15 to 2017-18 school years and bind it into a single data frame, using a single function (i.e., one line of code). Note, this may take a minute or two to run, depending on your Internet speed.
```{r}
# use function
ode_data <- map_dfr(15:18, download_file)

ode_data
```

Conduct some basic data cleaning to make your data file look like the following.

- Filter for only student groups coded as "White" or "Hispanic/Latino".

- Select variables related to the number of students in each of the levels (1 through 4), and not percentages or combined levels.

- Remove any row that has missing data in any of the n variables

```{r}
# clean data
ode_data_clean <- ode_data |>
  clean_names() |>
  filter(student_group == "White" | student_group=="Hispanic/Latino") |>
  select(academic_year, district, school, student_group, grade_level, number_level_1, number_level_2, number_level_3, number_level_4) |>
  pivot_longer(
    c(number_level_1, number_level_2, number_level_3, number_level_4),
    names_to = "level",
    names_prefix = "number_level_",
    values_to = "n",
    values_drop_na = TRUE
  )
  
ode_data_clean
```

### 2. (10 points)

Sum the n across grades and school for each district within each academic year, student group, and level. Calculate the number of schools (unique schools listed) for each district. Also move your data to a wider format where the student groups become new columns, filled by the district n. Remove districts that did not report numbers for one or both student groups (i.e., drop missing data from the new columns).

```{r}

school_counts <- ode_data_clean |>
  group_by(district) |>
  summarize(n_schools = n_distinct(school))

# I asked chatgpt about how, given the code below that I wrote (prior to the left_join), to add in the column n_schools and have it only be by district, not by every group. It said to create it separately then join, so I created it separately above and joined it at the bottom of this chunk. I tried not to look at the exact code it gave me so I could troubleshoot myself.
ode_sum <- ode_data_clean |>
  group_by(academic_year, district, student_group, level) |>
  summarise(n_sum = sum(n), .groups = "drop") |>
  pivot_wider(
    names_from = student_group,
    values_from = n_sum
  ) |> 
  clean_names() |>
  filter(!is.na(hispanic_latino) & !is.na(white)) |>
  left_join(school_counts) |>
  arrange(district)

ode_sum
```

## Part B: Achievement Disparities
```{r}
#remotes::install_github("datalorax/gapr")

library(gapr)
#ode_sum[1:4, ] |> # data source - just first four rows for first district
#  estimate_v("white", "hispanic_latino") # columns for reference/focal groups
```
### 1. (15 points)
Estimate the average achievement differences in terms of an effect size for every school district in the state that reported data on both student groups (i.e., using the data we created above), for each academic year. Conduct this analysis twice, once using group_by() |> nest() and map() and once using nest_by() |> summarize().
