---
title: "Midterm Test"
author: "Steven Todd McIntyre II"
cache: true
format: 
  html:
    code-fold: true
    embed-resources: true
---

## Install Dependencies

```{r}
library(tidyverse)
library(glue)
library(here)
library(rio)
library(janitor)
library(ggtext)
```
## Part A: Data

```{r}
# create function
download_file <- function(year) {
  link <- glue::glue("https://www.oregon.gov/ode/educator-resources/assessment/TestResults20{year}/pagr_schools_ela_raceethnicity_{year-1}{year}.xlsx")
  rio::import(link, setclass = "tibble", na = c("-", "--", "*"))
}
```

### 1. (10 points)

Use the function above to download all of the data from the 2014-15 to 2017-18 school years and bind it into a single data frame, using a single function (i.e., one line of code). Note, this may take a minute or two to run, depending on your Internet speed.
```{r}
# use function
ode_data <- map_dfr(15:18, download_file)

ode_data
```

Conduct some basic data cleaning to make your data file look like the following.

- Filter for only student groups coded as "White" or "Hispanic/Latino".

- Select variables related to the number of students in each of the levels (1 through 4), and not percentages or combined levels.

- Remove any row that has missing data in any of the n variables

```{r}
# clean data
ode_data_clean <- ode_data |>
  clean_names() |>
  filter(student_group == "White" | student_group=="Hispanic/Latino") |>
  select(academic_year, district, school, student_group, grade_level, number_level_1, number_level_2, number_level_3, number_level_4) |>
  pivot_longer(
    c(number_level_1, number_level_2, number_level_3, number_level_4),
    names_to = "level",
    names_prefix = "number_level_",
    values_to = "n",
    values_drop_na = TRUE
  )
  
ode_data_clean
```

### 2. (10 points)

Sum the n across grades and school for each district within each academic year, student group, and level. Calculate the number of schools (unique schools listed) for each district. Also move your data to a wider format where the student groups become new columns, filled by the district n. Remove districts that did not report numbers for one or both student groups (i.e., drop missing data from the new columns).

```{r}

school_counts <- ode_data_clean |>
  group_by(district) |>
  summarize(n_schools = n_distinct(school))

# I asked chatgpt about how, given the code below that I wrote (prior to the left_join), to add in the column n_schools and have it only be by district, not by every group. It said to create it separately then join, so I created it separately above and joined it at the bottom of this chunk. I tried not to look at the exact code it gave me so I could troubleshoot myself.
ode_sum <- ode_data_clean |>
  group_by(academic_year, district, student_group, level) |>
  summarise(n_sum = sum(n), .groups = "drop") |>
  pivot_wider(
    names_from = student_group,
    values_from = n_sum
  ) |> 
  clean_names() |>
  filter(!is.na(hispanic_latino) & !is.na(white)) |>
  left_join(school_counts) |>
  arrange(district)

ode_sum
```

## Part B: Achievement Disparities
```{r}
#remotes::install_github("datalorax/gapr")

library(gapr)
#ode_sum[1:4, ] |> # data source - just first four rows for first district
#  estimate_v("white", "hispanic_latino") # columns for reference/focal groups
```
### 1. (15 points)
Estimate the average achievement differences in terms of an effect size for every school district in the state that reported data on both student groups (i.e., using the data we created above), for each academic year. Conduct this analysis twice, once using group_by() |> nest() and map() and once using nest_by() |> summarize().

```{r}
# this is my original code, which didn't work because, for some reason, estimate_v is failing on a few items

#analysis1 <- ode_sum |>
#  group_by(academic_year, district) |>
#  nest() |>
#  mutate(
#    v_estimate = map(data, ~estimate_v(.x, "white", "hispanic_latino")),
#    v = map_dbl(v_estimate, "v")
#  )

# I pasted the above code into chatGPT and asked it how I can just remove those rows where the analysis fails, since analysis2 just seems to be automatically removing them. 

# The precise prompt was [pasted code] + "The above is using estimate_v to do an analysis on each district by academic year, but whenever estimate_v returns an empty data frame, v can't be determined and it throws an error..."

# after it replied, I added "what if i want to just remove the empty rows"

# it gave me the following:

analysis1 <- ode_sum |>
  group_by(academic_year, district) |>
  nest() |>
  mutate(
    v_estimate = map(data, ~estimate_v(.x, "white", "hispanic_latino"))
  ) |>
  filter(map_lgl(v_estimate, ~nrow(.x) > 0 && "v" %in% names(.x))) |>
  mutate(
    v = map_dbl(v_estimate, ~.x$v[1])
  )
```
```{r}
# nest_by + summarize
analysis2 <- ode_sum |>
  ungroup() |>
  nest_by(academic_year, district) |>
  summarize(
    data = list(data),
    v_estimate = list(estimate_v(data, "white", "hispanic_latino") ),
    # had to ask a similar question of chatgpt here as in analysis 1
    v = {
      if (nrow(v_estimate) == 0) NA_real_
      else v_estimate[["v"]]
    }
  )  |> 
  filter (!is.na(v))
```

### 2. (10 points)

Reproduce the following plot to confirm that the estimates are equivalent across methods.

```{r}
analysis1 |>
  mutate(v1 = v) |>
  select(district, v1) |>
  left_join(select(mutate(analysis2, v2 = v), district, v2)) |>
  ggplot(aes(x = v2, y = v1)) +
    geom_point(size = 4, color = "grey") +
    geom_smooth(method = lm, color = "cyan", size = 1) +
    theme_minimal() +
    labs(
      title = "Method Equivalence Test",
      x = "V (nest_by) > summarize) )",
      y = "V ('group_by() |> nest() |> mutate() |> map()')"
    )
```

### 3. (15 points)

The plot below shows the achievement difference estimates for one school district across years. Produce a similar plot to the below for each of the first 100 school districts. Once you’ve created all the plots, reproduce them again using an alternative method. Make sure you don’t miss the caption noting the number of schools in the district.

```{r}
dists <- unique(analysis2$district)[1:100]

analysisprime <- analysis2 |>
  filter(district %in% dists) |>
  summarize(
    v = v,
    n_schools = data[[1]]["n_schools"][[1]][[1]],
    v_se = v_estimate[[1]][["v_se"]][[1]]
  )

# method 1: group_by, nest, mutate, map2
ode_plots1 <- analysisprime |>
  ungroup() |>
  group_by(district) |>
  nest() |>
  mutate(
    plot = map2(data, district, 
                ~.x |>
                  ggplot(aes(x=academic_year)) +
                    # asked gemini how to make the next line "how to make
                    # error bars for a geom point based on an error value"
                    theme_minimal() +
                    geom_hline(yintercept = 0, color = "darkred") +
                    geom_line(aes(y=v), color = "lightgrey", group = 1, size = 0.5) +
                    geom_errorbar(aes(ymin = v - v_se, ymax = v + v_se), color = "darkgrey") +
                    geom_point(aes(y = v, color = v>0)) +
                    labs(
                      title = paste0("Achievement Gap Estimates: ", .y),
                      subtitle = "Negative values indicate <span style='color:salmon;'>White</span> students have higher average achievement\n\nPositive values indicate <span style='color:#00BFC4;'>Hispanic/Latino</span> students have higher average achievement",
                      x = "School year",
                      y = "Average achievement difference (SD units)",
                      caption = paste0(
                        "Note: District included ", 
                        .x$n_schools[[1]], 
                        " schools.")
                    ) +
                    theme(
                      plot.subtitle = element_markdown(size=10),
                      legend.position = "none"
                    )
    )
  )

ode_plots1$plot[[7]]

# method 2: nest_by()
ode_plots2 <- analysisprime |>
  ungroup() |>
  nest_by(district) |>
  mutate(
    plot = list(ggplot(data, aes(x=academic_year)) +
                  # asked gemini how to make the next line "how to make
                  # error bars for a geom point based on an error value"
                  theme_minimal() +
                  geom_hline(yintercept = 0, color = "darkred") +
                  geom_line(aes(y=v), color = "lightgrey", group = 1, size = 0.5) +
                  geom_errorbar(aes(ymin = v - v_se, ymax = v + v_se), color="darkgrey") +
                  geom_point(aes(y = v, color = v>0)) +
                  labs(
                    title = paste0("Achievement Gap Estimates: ", district),
                    subtitle = "Negative values indicate <span style='color:salmon;'>White</span> students have higher average achievement\n\nPositive values indicate <span style='color:#00BFC4;'>Hispanic/Latino</span> students have higher average achievement",
                    x = "School year",
                    y = "Average achievement difference (SD units)",
                    caption = paste0(
                      "Note: District included ", 
                      data$n_schools[[1]], 
                      " schools.")
                  ) +
                  theme(
                    plot.subtitle = element_markdown(size=10),
                    legend.position = "none"
                  )
    )
  )

ode_plots2$plot[[7]]
```

### 4. (10 points)

Save the plots into a “plots” directory. Make sure the file names are meaningful.

```{r}


ode_plots1 |>
  ungroup() |>
  rowwise() |>
  summarize(
    ggsave(
      here(paste0(
        "plots/v1/", 
        str_replace_all(
          tolower(district), " ", "-"
        ),
        "-achievement-gap-v1.png")
      ),
      plot,
      width = 9.5, 
      height = 6.5,
      dpi = 500
    )
  )

ode_plots2 |>
  ungroup() |>
  rowwise() |>
  summarize(
    ggsave(
      here(paste0(
        "plots/v2/", 
        str_replace_all(
          tolower(district), " ", "-"
        ),
        "-achievement-gap-v2.png")
      ),
      plot,
      width = 9.5, 
      height = 6.5,
      dpi = 500
    )
  )
```

